Benchmarking neuromorphic systems with Nengo

### 2015-09-13 paper/rebuttals.txt

Reviewer 1
==========

1. What are the main findings reported in this manuscript?
----------------------------------------------------------

> The main contribution of this work is a benchmarking suite designed
> to test neuron simulators in a relevant and objective way. This is
> performed by analyzing three fundamental metrics - compliance,
> accuracy, and speed - under different network test
> conditions. Compliance metrics determine feature coverage for a
> neural simulator, while the accuracy and speed metrics are tested
> under four neurally-relevant operations: representation with
> robustness to noise, nonlinear computation, stability with recurrent
> dynamics, and large-scale interaction of multiple recurrent
> populations, all implemented with neurons. Finally, five neural
> simulators with different algorithmic and hardware optimizations are
> tested on these benchmarks.

@@

2. Other comments on the introduction.
--------------------------------------

> The introduction is well-written and motivates the study; the
> background provides sufficient detail to understand the architecture
> of this solution.

@@

3. Other comments on the materials and methods.
-----------------------------------------------

> The materials and methods describes well the metrics and models used
> to evaluate the backends.

@@

4. Other comments on the results.
---------------------------------

> 1. Further clarification should be given as to why the SpiNNaker
> backend was not tested for compliance, as lines 276-77 are not
> sufficient. The reason is not clear as build and run time for the
> examined models in Figures 5-7 do not seem prohibitive to running
> 212 tests; additionally, SpiNNaker supports a wide variety of neuron
> models and parameters, so it appears adequate to support the
> necessary descriptions required by Nengo. Please clarify.

You are completely correct that this comment is not sufficient.
A technical answer would be that SpiNNaker has a lot of
set up and tear down code, which makes running the test suite
slow and error-prone. The tests creates a large number of
small models and runs them for small amounts of time;
this results in the SpiNNaker boards' memory filling up with
improperly freed resources. So, the real answer is more a
time management one; we decided that getting the unit tests
running would take too much time for this paper.
However, this answer is not very satisfying, so we have
put in the time to run the test suite with the SpiNNaker
backend. Table 1 (now Table 2) now also gives results
for the SpiNNaker backend, and the text has been updated
accordingly.

> 2. Lines 320-325 are not clear.

> a) While Figure 6 does show an increase for the Nengo reference
> backend run times by removing passthrough nodes, the run time appears
> to be a 15% increase (~7x vs ~6x slower than real-time, Figure 6
> green, BG Sequence to BG Sequence *), which is approximately the same
> relative speedup as that offered by removing probes for the SpiNNaker
> backend (~1.5x to ~1.2x, Figures 6 and 7, BG sequence, blue). So,
> "significant" is a strong claim and should be clarified.

Good point. We used 'significant' colloquially here and
have reworded this paragraph; it now reads:

  While passthrough nodes were removed in the basal ganglia sequence
  model for improved accuracy, it is interesting to note that speed is
  also impacted in this version of the model.  While the distilled and
  Brainstorm backends are unaffected, both the SpiNNaker and OpenCL
  backends run faster when passthrough nodes are removed.  This result
  indicates that passthrough nodes contribute to SpiNNaker's overhead.
  The reference backend, on the other hand, runs slower, indicating
  that the computational costs of the additional connections
  introduced when removing passthrough nodes are greater than the
  costs of the passthrough nodes.

> b) Additionally, Figure 7 seems to indicate that removing probes
> actually makes the network slower for the reference backend for BG
> sequence * (~8x vs ~6.8x, Figures 6 and 7, light blue, BG Sequence
> *). Either outliers / standard deviations are stronger than they
> appear from the plot, or the reason for the network slowing down with
> probes removed should be explained.

We did not notice the slowdown upon initial submission,
so seeing this was puzzling to us as well.
To investigate whether this effect was real,
we first replicated it on two additional machines;
the version without probes was significantly slower
than the version with probes on all machines,
as measured by non-overlapping
bootstrapped 95% confidence intervals.
Then, we investigated the internal structures of the two
versions of networks, in terms of the step functions
produced by them.
The version with probes had 600 step functions;
595 of these implemented the same equations
as the networks without probes, while the
remaining five functions implemented the probing.
The simulator loop also indicated that
the version without probes should have executed
strictly fewer operations than the version with probes.
We conclude that the reason for the speedup
must have been somehow related to
a lower level of abstraction than we investigated
(e.g., the data was better organized in RAM
in the version with probes).

Very few networks in Nengo have zero probes,
so there were some obvious opportunities to optimize
the reference backend for the special case
in which no probes exist. After implementing
two such optimizations, we were able to consistently
achieve identical performance with probes and without.
However, it was not our intention to perform
a detailed comparison of the affect of probes
on runtimes, and we feel that including Figure 7
(which only had a few sentences of text associated with it)
was more distracting than illuminating.
For these reasons, we have removed Figure 7 in this revision.

> c) Finally, the speedup by removing passthrough nodes for the OpenCL
> backend (Figure 6, green, BG sequence to BG sequence*) disappears once
> probes are removed (Figure 7, green, BG sequence to BG sequence*).

As discussed above, Figure 7 has now been removed.
We believe that, with more robust benchmarks
and investigation of the probe system in the OpenCL backend,
the speedup would still occur,
but will address this in future work that focuses
on probing across multiple backends.

> Essentially, please verify that the error bars on the plots are
> representative of the true confidence intervals. If they are correct,
> there seem to be some anomalies in the data (removing reporting output
> slows down the reference backend for BG*, speedups in OpenCL become
> slowdowns once reporting output is removed). If this all has to do
> with build time / overhead / short runtimes / network latency, then
> that should be mentioned in the text.

We have verified that the error bars reported are bootstrapped 95%
confidence intervals, as claimed in the figure captions of
Figures 5 and 6. As previously mentioned, we also replicated
the paradoxical slowdown with probes removed; however,
with the removal of Figure 7, we believe that additional text attesting
to the validity of the confidence intervals is not warranted.

> 3. In Figure 3, the meaning of negative frequencies (-1 Hz, -2 Hz)
> should be at least briefly mentioned. Presumably, since this is a
> circular oscillator, it is being run in reverse. However, it's not
> clear to me why negative frequencies should be at all different than
> positive frequencies, and how the range -2 Hz to 2 Hz was chosen. A
> citation that references this range in biology for CPGs, for example,
> would bolster this choice.

The meaning of negative frequencies has been made explicit in text;
line 239 now notes that "negative frequencies indicate
oscillations in the reverse direction." In that line, we have
also added a footnote with a citation to Duysens and Van de Crommert, 1998,
which provides some evidence that locomotive CPGs may run in reverse
for backward locomotion.

In terms of why the range -2 to 2 was chosen,
this was because we have used oscillators of these frequencies
in previous model. In line 219 we note that
"We test each model with parameters typically used
in large-scale models, but make these parameterized models
available ... for those who wish to explore additional cases."
The parameter is trivial to change, up to an upper frequency limit
which depends on the properties of the neurons in use;
we would be happy to show results for a different set of frequencies
if that would be preferable.

> 4. Table 1, the compliance of the four backends, should be explained
> more fully in the caption. The levels of coverage are surprisingly low
> given that all backends are able to perform the main cognitive-level
> tasks; common failure cases (e.g., missing neuron models) should be
> mentioned to show why these are not necessary.

The caption for the compliance table (which now includes all 5 backends)
has been updated. It now reads:

  Compliance of the five backends. Compliance is the number of
  functional tests passed by a particular backend. We have divided
  these tests into tests parameterized by neuron type, and those not
  parameterized.  Non-parameterized tests are either designed to work
  with only one neuron type, or use leaky integreate-and-fire (LIF)
  neurons, which are implemented on all backends. For parameterized
  tests, we have separated the compliance for those tests using LIF
  neurons, which all backends implement, and for any additional neuron
  types implemented. Since there are 31 parameterized tests, the
  maximum compliance is 31 times the number of neuron types
  implemented.  The reference backend implements three additional
  neuron types, while OpenCL, Distilled, and Brainstorm implement one
  additional neuron type, and SpiNNaker only implements LIF neurons.

Additionally, a paragraph explaining common failure cases
is included in the main text in lines 311-321. It reads:

  It should be noted that these test suites are being run on the
  distilled, Brainstorm, and SpiNNaker backends for the first time in
  this study.  In other words, these results are the first objective
  interrogation of each backend's feature set.  In many cases, tests
  failed by backends are tests of features recently added to Nengo,
  such as stochastic processes for injecting current noise.  As such,
  we expect compliance on all backends to rise quickly as backend
  developers implement these new features.  The OpenCL backend has
  significantly higher compliance in large part because it is
  developed by the same group that develops the reference backend.
  However, there are some notable features missing from some backends,
  such as learning through plasticity rules applied to
  neuron-to-neuron connections, and implementation differences, such
  as one timestep delays on connections with no synaptic filter, that
  may remain even when backends are brought up to date with the
  reference implementation.  It is likely that tests will be rewritten
  in the future to allow some implementation differences if they do
  not affect simulation accuracy.

@@

5. Other comments on the discussion or the conclusions.
-------------------------------------------------------

> The initially stated conclusion that significant speedups can be
> obtained with specialized hardware without sacrificing accuracy is
> good, although perhaps not very surprising. However, the greater
> challenge of creating a list of cognitively-relevant tasks and an
> infrastructure for objective testing is significant. Moreover, the
> improved accuracy of SpiNNaker with passthrough nodes removed,
> discovered by these tests, is an encouraging sign that the
> optimization loop of hardware and implementation through this test
> framework can yield real improvements.
>
> The biggest omission from this discussion and the manuscript is any
> mention of efficiency. Neuromorphic systems have long held that
> their primary advantage is significant efficiency, and the lack of
> ability of the benchmarking to capture that is unfortunate. This
> could perhaps be implemented as an optional query on the backend;
> the power consumption of most CPUs and GPUs can be publicly
> referenced and approximated, or the implementation could provide its
> own statistics for specialized hardware. Perhaps the Brainstorm
> backend’s main advantage is in power efficiency, and there should be
> resulting statistics that can show that. Regardless of
> implementation, the authors should at least mention that efficiency
> could be evaluated in future work.

This is a great point, and indeed points out one of the weaknesses
of the "standardization" approach that we are espousing with Nengo:
new metrics may take a long time to be implemented since they
must be implemented on all backends.

We have a paragraph to the discussion section
to discuss this issue, and to register our desire to collect
efficiency information in future work. Lines 417-423 now read:

  One inherent weakness of using Nengo as a standard platform for
  benchmarking neuromorphic systems is that new benchmarking
  capabilities may take a long time to be standardized and developed.
  If one wishes to add a new metric, such as power consumption, we
  must first come to a consensus on a suitable interface to this
  information through Nengo.  Once consensus is reached, it must be
  implemented and tested on all backends before benchmarks can be
  written using that quantity.  Despite this limitation, we believe
  that it is possible to use Nengo to collect power consumption
  information, and plan to implement efficiency comparisons in future
  work.

@@

6. Please add here any further comments on this manuscript.
--------------------------------------------------------

> This work addresses a very challenging topic - that is, how to
> objectively compare systems designed for different objectives in a
> high-dimensional benchmarking space. The approach of choosing
> cognitive tasks and measuring the ability of neuromorphic systems to
> perform them is insightful and relevant.

> It does, however, represent only an initial step towards obtaining
> the necessary concrete metrics to measure neuromorphic systems. Many
> quantities remain untested in this framework; researchers who build
> neuromorphic systems designed for neural simulation, the target
> audience of this manuscript, focus on optimizing many of the
> following quantities: simulation speed, simulation accuracy, energy
> per synaptic event, accurate timing, number of neurons represented,
> number of connections represented, repertoire of neural models,
> online learning, and likely others. Of these, the test framework
> objectively compares only along the first two, though the compliance
> tests likely would stress some of the others. I hope to see future
> work along these lines to expand the reach of these tests.

> Nonetheless, this work provides a substantive proposal to begin
> evaluating neuromorphic neural simulators.

@@

Reviewer 2
==========

1. What are the main findings reported in this manuscript?
----------------------------------------------------------

> This paper describes how the Nengo test suite can be used for
> benchmarking neuromorphic systems and compare them to neural
> simulators on general-purpose computers. It introduces the Nengo
> test suite, describes three benchmark metrics (compliance, accuracy,
> speed) and shows benchmark results for four tests on five different
> backends. Of these backends, only one is actually running on
> neuromorphic hardware, one is a simulator of a neuromorphic system,
> and three are software simulators (with one employing
> GPUs). Benchmark results are shortly discussed.

@@

2. Does the introduction present the study in an appropriate context?
---------------------------------------------------------------------

> No

> The introduction gives a clear motivation for the work in the
> article, but it does not state which works have been done previously
> on that topic. For example, there are several articles investigating
> accuracy in neuromorphic hardware for specific benchmark
> problems. As three of the backends are software-based, it would also
> be appropriate here to cite prior work on accuracy/speed in neural
> software simulators.

We have added two paragraphs to the introduction citing prior work
benchmarking SpiNNaker, FACETS, and Brian,
and situating this study in that context in lines 19-35.
They read:

  Prior work benchmarking neural simulators and neuromorphic hardware
  has focused on low-level neural performance.  For example,
  \citet{sharp2013} showed that SpiNNaker can simulate a recurrent
  network of leaky integrate-and-fire neurons with similar firing rate
  and inter-spike intervals as the NEST neural simulator, but around
  six times faster.  \citet{stromatias2013} showed that SpiNNaker's
  power consumption varies between 15 and 37 Watts (0.5--0.8 Watts per
  chip) depending on the number of neurons being simulated.
  \citet{goodman2008brian} showed that Brian simulated a randomly
  connected network of 4000 leaky integrate-and-fire twice as fast as
  an equivalent Matlab implementation, but around three times as slow
  as a C implementation.  In all of these cases, none of the networks
  elicited activity that could be directly related to experimentally
  recorded data on a behavioral task.

  In contrast, \citet{ehrlich2010} and \citet{bruderle2011} have
  presented a set of benchmarks that target the FACETS neuromorphic
  system through the PyNN Python package.  These benchmarks include an
  attractor-based memory model, a model of self-sustained AI states,
  and a Synfire Chain, all of which are directly related to
  neuroscientific experiments.  We aim to build on this line of
  research to provide an unbiased, efficient, and future-proof set of
  benchmarks that focuses on high-level functional performance using
  Nengo instead of PyNN.  We have previously shown that Nengo is an
  order of magnitude faster than the simulators that PyNN targets
  \citep{bekolay2013}, and have recently implemented backends that
  target neuromorphic hardware.

@@

3. Other comments on the introduction.
--------------------------------------

> As the Nengo test suite was not designed for benchmarking, it would
> be helpful to have some indications on the limitations of the method
> (its potential is described already pretty clear) - this could be
> done in the Discussion section as well.

We have added a paragraph to the discussion noting that Nengo's
standards-based development and frontend-backend separation mean that
development can be slow (e.g., collecting power consumption
would mean figuring out a standard for this and then implementing
it across all backends) in lines 417-423. It reads:

  One inherent weakness of using Nengo as a standard platform for
  benchmarking neuromorphic systems is that new benchmarking
  capabilities may take a long time to be standardized and developed.
  If one wishes to add a new metric, such as power consumption (as was
  done in \citealp{stromatias2013}), we must first come to a consensus
  on a suitable interface to this information through Nengo.  Once
  consensus is reached, it must be implemented and tested on all
  backends before benchmarks can be written using that quantity.
  Despite this limitation, we believe that it is possible to use Nengo
  to collect power consumption information, and plan to implement
  energy efficiency comparisons in future work.

@@

4. Other comments on the materials and methods.
-----------------------------------------------

> I found the separation in Background and Methods helpful for
> understanding, even if it does not follow the standard article
> structure. Some points should be clarified:
>
> - The benchmarks make use of current input to neurons. However,
>   quite a lot of neuromorphic systems do only offer spike input, or
>   current inputs are not so flexible to program. Of course, current
>   input can be emulated by spike input, but that may add some
>   additional noise/inaccuracy. This should be dicsussed in the
>   Background/Methods sections; also, an indication on how the
>   currents look like in the benchmark models would be helpful, so
>   that designers of neuromorphic systems get an idea of what would
>   be required.

This is an excellent point that we really should have addressed
in the initial submission. The paper that we cited in the section
on the SpiNNaker backend (Mundy et al., 2015) gives a detailed
account of how SpiNNaker uses decoded values to change the
input currents to individual neurons, and a prior publication
(Galluppi et al., 2012, which is now cited) describes a process
to implement Nengo with primarily spike-based communication
on SpiNNaker. We have added a paragraph to the Methods section
with brief descriptions of how SpiNNaker handles currents,
and how other neuromorphic hardware can be made to handle
currents from Nengo as well in lines 285-297.
The paragraph reads:

  Unlike software backends, considerable effort is taken to translate
  a Nengo model into something that can run on a SpiNNaker board.
  Fortunately, SpiNNaker can be reprogrammed as it is composed of a
  large collection of chips, each with 18 ARM processing cores.  In
  particular, this allows SpiNNaker to take advantage of Nengo's
  encoding and decoding capabilities, greatly reducing the amount of
  RAM needed \citep{mundy2015}.  The SpiNNaker backends accomplishes
  this by using decoded values to determine changes to the input
  currents to each neuron, rather than using purely spike-based
  transmission.  However, a previous implementation of a Nengo
  SpiNNaker backend \citep{galluppi2012} used primarily spike-based
  communication, only making an explicit signal-to-spike encoding at
  the beginning of the network and an explicit spike-to-signal
  decoding at the end of a network.  Neuromorphic hardware that can
  only communicate through spikes can be made to interact with Nengo
  through explicit encoding and decoding processes like the one used
  in \citealp{galluppi2012}, or by using Nengo on the host machine to
  explicitly generate spike patterns that are communicated to the
  device.  However, these methods may introduce additional noise in
  the simulation, resulting in inaccuracies.

> - At several points, the authors mention that seeds can be set
>   explicitely for deterministic experiment runs. Is there a reason
>   for this? For benchmarking, setting a fixed seed is misleading in
>   my opinion, as it suggests that all backends run exactly the same
>   experiment. However, as the order in which the single parts of a
>   network are generated is not the same, this is not the case in
>   general. Furthermore, neuromorphic hardware systems may use
>   built-in noise sources (i.e. true random numbers). All that
>   requires a statistical comparison of results over several
>   experiment repetitions, as the authors do in their benchmarks.

### 2015-09-13 paper/paper.tex
Edits in response to reviewer feedback

Benchmarking is a notoriously difficult task.
Benchmarks are often created by
the creators of the tools being benchmarked,
resulting in biased comparisons favoring their tool.
Benchmarking can be an inefficient process,
as the tool being benchmarked
may need changes to collect
certain performance metrics.
Even once that effort is undertaken,
benchmarks are often run a few times
and then forgotten,
quickly becoming obsolete.
We propose that the Nengo test suite
can serve as an unbiased, efficient, and future-proof
set of benchmarks for neuromorphic systems.

Nengo is a rigorously tested software package
for building and simulating
large-scale neural models
that can perform cognitively relevant tasks.
It provides a high-level API (frontend)
that can express large-scale models concisely
and in a platform-independent manner.
Several Nengo-compatible simulators (backends)
have been developed that can
run Nengo models on diverse platforms,
including neuromorphic hardware.
Nengo's separation between frontend and backend,
as well as its extensive test suite,
provide standardized comparisons
between different neuromorphic implementations
with respect to functional performance.

@@

There are two key features of Nengo
(from version 2.0 onward)
that enable rapid benchmarking of neuromorphic systems.
The first is a decoupling of model creation and simulation,
resulting in a platform independent frontend interface
to any backend that implements a certain set of requirements.
The second is a ``fuzzy'' test suite that ensures
Nengo can be used to create models that solve cognitive tasks.
The test suite makes ample use of test fixtures
to enable data collection while testing on arbitrary backends.

@@

Finally, the \textit{Network} is a
container for the other four objects.
The only parameter on the network
is an optional integer seed;
setting this should make
all of the random factors
in the model deterministic.
Aside from managing the random seed,
the network is responsible for maintaining
a network-specific set of default parameters
for the four other objects.
This results in shorter and less error-prone
model creation scripts.

@@

The first method is \texttt{\_\_init\_\_},
which is a special Python method for initializing objects.
This method must accept a network as its first argument.
It can then accept optional arguments depending on
the capabilities and requirements of the backend;
the reference backend, for example,
accepts \texttt{dt},
which is the length of each timestep.
The purpose of \texttt{\_\_init\_\_} is to
set up the low-level system
that implements the high-level objects
contained in the provided network.
In the reference backend,
this involves sampling from the distributions
in all of the parameters of all objects,
solving for decoding weights and connection weights,
setting up data structures for probed data,
and so on.
When parts of the model are randomly sampled,
the random factors should become deterministic
when a network's seed is set.

@@

\subsection{Fuzzy testing}

@@

Nengo's test suite defines several test fixtures
to support fuzzy testing
that collects and visualizes performance metrics
on multiple backends.

\paragraph{\texttt{rng} \& \texttt{seed}}
enable deterministic testing despite
all of the random factors that affect a model.
Test functions that generate random values
(as inputs to a model, for example)
generate those values using \texttt{rng},
which is a random number generator
with an interface identical to
NumPy's random module.
When a test function creates a network,
it passes in \texttt{seed} to make
the random components
of the frontend-to-backend translation deterministic.
Importantly, these random components
are seeded by hashing a string uniquely
identifying each test function,
ensuring that tolerances must be set
wide enough that changing the seed
should only rarely result in test failures.
The hash function is also seeded
with a global integer,
which is changed periodically to ensure
that tests have good tolerances.

@@

\paragraph{3. Controlled oscillator.}
In this model,
a three-dimensional ensemble is recurrently connected
such that the first two dimensions continuously
traverse a limit cycle
(i.e., they implement a cyclic attractor),
and the third dimension controls the speed
and direction of the oscillation.
As input, we provide an initial stimulus
to start the oscillation,
and provide a control signal
to cause oscillation at
2~Hz, 1~Hz, 0~Hz, -1~Hz, and -2~Hz.
In total, the model is comprised of 600 LIF neurons.
This model tests each backend's ability
to stably implement a dynamical system,
which is required for many cognitive functions
like working memory and motor control.

@@

@@

@@

The versions of software
used for benchmarking can be found in
Table~\ref{Tab:02}.

@@

\subsection{Compliance}

Table~\ref{Tab:01} gives compliance results for
all backends except the SpiNNaker backend,
which is not yet set up to run the Nengo test suite.
Of the four currently running unit tests,
the OpenCL backend has the highest compliance
(other than the reference backend).
The distilled backend has lower compliance
than OpenCL, which follows from its stated goal
of being simple but not necessarily feature-rich.
The Brainstorm backend is the least compliant,
but it is still an early experimental test bed;
this is not necessarily indicative of the
final Brainstorm hardware backend being
less compliant than other backends.

@@

Figure~\ref{fig:01} gives accuracy results
for each backend on the chained communication channel model.
The boxplot shows that all five backends can
implement this model accurately,
despite five layers of processing
that each introduce noise.

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
is less accurate than the other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
The median accuracy of the SpiNNaker is
the same as the distilled and Brainstorm backends,
but there are a high proportion of model instances
with less accuracy than the median,
as compared to other backends.

Figure~\ref{fig:04} gives accuracy results
for the basal ganglia sequence model.
For the version of the model with passthrough nodes,
all of the backends perform similarly,
except for the SpiNNaker backend.
All other backends have a transition time
around 43 milliseconds,
but the SpiNNaker backend has a median transition time
around 51 milliseconds;
its interquartile range
is also significantly larger
than that of other backends.
However, its performance for
the version of the model
with passthrough nodes removed
is a closer match to the other backends.
Its median transition time
is around 47 milliseconds,
and its interquartile range
is indistinguishable from the other backends.

@@

Interestingly, while removing passthrough nodes
results in significantly slower run times
for the reference backend,
it speeds up the OpenCL and SpiNNaker backends slightly.
This result indicates that passthrough nodes
add to the overhead incurred by
SpiNNaker and OpenCL.
Other backends also incur overhead,
especially when data is recorded
through probes.
We investigated the possibility
that the overhead involved in probes
dominated the run time for some backends.
However, we found that
removing probes from these models
made very little impact on
the total time taken in a \texttt{run} call
(see Figure~\ref{fig:07}).

>>>>

Benchmarking is a notoriously difficult task.
Benchmarks are often created by
the creators of the tools being benchmarked,
resulting in biased comparisons favoring their tool.
Benchmarking can be an inefficient process,
as the tool being benchmarked
may need changes to collect
certain performance metrics.
Even once that effort is undertaken,
benchmarks are often run a few times
and then forgotten,
quickly becoming obsolete.
However, benchmarks can spur progress
as tool developers have an objective
metric to maximize or minimize.

Prior work benchmarking neural simulators
and neuromorphic hardware has focused
on low-level neural performance.
For example, \citet{sharp2013}
showed that SpiNNaker can simulate
a recurrent network of leaky integrate-and-fire neurons
with similar firing rate and inter-spike intervals
as the NEST neural simulator,
but around six times faster.
\citet{stromatias2013} showed that
SpiNNaker's power consumption varies
between 15 and 37 Watts
(0.5--0.8 Watts per chip)
depending on the number of neurons
being simulated.
\citet{goodman2008brian} showed that
Brian simulated a randomly connected
network of 4000 leaky integrate-and-fire
twice as fast as an equivalent Matlab implementation,
but around three times as slow as
a C implementation.
In all of these cases,
none of the networks
elicited activity that could be
directly related to experimentally recorded data
on a behavioral task.

In contrast,
\citet{ehrlich2010} and \citet{bruderle2011}
have presented a set of benchmarks that target
the FACETS neuromorphic system
through the PyNN Python package.
These benchmarks include
an attractor-based memory model,
a model of self-sustained AI states,
and a Synfire Chain,
all of which are
directly related to neuroscientific experiments.
We aim to build on this line of research
and provide an unbiased,
efficient, and future-proof set
of benchmarks that focuses
on high-level functional performance
using Nengo instead of PyNN.
We have previously shown that Nengo
is an order of magnitude faster than
the software simulators that PyNN targets
\citep{bekolay2013},
and have recently implemented backends
that target neuromorphic hardware.

In this study,
we propose that the Nengo test suite
can serve as an unbiased, efficient, and future-proof
set of benchmarks for neuromorphic systems.
Nengo is a rigorously tested software package
for building and simulating
large-scale neural models
that can perform cognitively relevant tasks.
It provides a high-level API (frontend)
that can express large-scale models concisely
and in a platform-independent manner.
Several Nengo-compatible simulators (backends)
have been developed that can
run Nengo models on diverse platforms,
including neuromorphic hardware.
Nengo's separation between frontend and backend,
as well as its extensive test suite,
provide standardized comparisons
between different neuromorphic implementations
with respect to functional performance.

@@

There are two key features of Nengo
(from version 2.0 onward)
that enable rapid benchmarking of neuromorphic systems.
The first is a decoupling of model creation and simulation,
resulting in a platform independent frontend interface
to any backend that implements a certain set of requirements.
The second is a functional test suite that ensures
Nengo can be used to create models that solve cognitive tasks.
The test suite makes ample use of test fixtures
to enable data collection while testing on arbitrary backends.

@@

Finally, the \textit{Network} is a
container for the other four objects.
The only parameter on the network
is an optional integer seed;
setting this should make
all of the random factors
in the model deterministic,
which is important for testing
and debugging models,
but is not used for benchmarking.
The network is also responsible for maintaining
a network-specific set of default parameters
for the four other objects.
This results in shorter and less error-prone
model creation scripts.

@@

The first method is \texttt{\_\_init\_\_},
which is a special Python method for initializing objects.
This method must accept a network as its first argument.
It can then accept optional arguments depending on
the capabilities and requirements of the backend;
the reference backend, for example,
accepts \texttt{dt},
which is the length of each timestep.
The purpose of \texttt{\_\_init\_\_} is to
set up the low-level system
that implements the high-level objects
contained in the provided network.
In the reference backend,
this involves sampling from the distributions
in all of the parameters of all objects,
solving for decoding weights and connection weights,
setting up data structures for probed data,
and so on.

@@

\subsection{Functional testing}

@@

Nengo's test suite defines several test fixtures
to support collecting and visualizing performance metrics
on multiple backends.

@@

\paragraph{3. Controlled oscillator.}
In this model,
a three-dimensional ensemble is recurrently connected
such that the first two dimensions continuously
traverse a limit cycle
(i.e., they implement a cyclic attractor),
and the third dimension controls the speed
and direction of the oscillation.
As input, we provide an initial stimulus
to start the oscillation,
and provide a control signal
to cause oscillation at
2~Hz, 1~Hz, 0~Hz, -1~Hz, and -2~Hz.
Negative frequencies indicate oscillations
in the reverse direction.\footnote{There is some evidence
  that neural oscillators such as locomotive central pattern generators
  accomplish forward and backward locomotion
  with the same neural mechanism \citep{duysens1998}.
  While this does not mean that all neural oscillators must
  be able to operate in the reverse direction,
  we believe that the capability to traverse
  the limit cycle forward and backward is advantageous,
  and in the current design requires no additional neural resources.}
In total, the model is comprised of 600 LIF neurons.
This model tests each backend's ability
to stably implement a dynamical system,
which is required for many cognitive functions
like working memory and motor control.

@@

We have used relatively small
models run for short times in order
to run many iterations on all backends.
However, all of these models
could be made significantly larger
by increasing the number of neurons used
(i.e., adjusting the \texttt{n\_neurons}
parameter on ensembles),
and increasing the dimensionality
of the signals represented in the model.
Additionally, existing tests
of the semantic pointer architecture
within Nengo use significantly more neurons
than the models presented here,
and could be adapted into benchmarks.

@@

Unlike software backends,
considerable effort is taken
to translate a Nengo model into
something that can run on a SpiNNaker board.
Fortunately, SpiNNaker can be reprogrammed
as it is composed of a large collection
of chips, each with 18 ARM processing cores.
In particular, this allows SpiNNaker
to take advantage of Nengo's
encoding and decoding capabilities,
greatly reducing the amount of RAM needed
\citep{mundy2015}.
The SpiNNaker backends accomplishes this
by using decoded values to determine
changes to the input currents to each neuron,
rather than using purely
spike-based transmission.
However, a previous implementation
of a Nengo SpiNNaker backend
\citep{galluppi2012}
used primarily spike-based communication,
only making an explicit signal-to-spike
encoding at the beginning
of the network and an explicit
spike-to-signal decoding
at the end of a network.
Neuromorphic hardware that
can only communicate through spikes
can be made to interact with Nengo
through explicit encoding and decoding processes
like the one used in \citealp{galluppi2012},
or by using Nengo on the host machine
to explicitly generate spike patterns
that are communicated to the device.
However, these methods may introduce
additional noise in the simulation,
resulting in inaccuracies.

@@

The versions of software
used for benchmarking can be found in
Table~\ref{Tab:01}.

@@

\subsection{Compliance}
Table~\ref{Tab:02} gives compliance results for all backends.
Compliance has been separated into tests
that use a backend's default neuron type
(leaky integrate-and-fire in all of the backends tested),
and tests that are parameterized by neuron type.
Of the non-reference backends,
the OpenCL backend has the highest compliance.
The distilled backend has lower compliance
than OpenCL, which follows from its stated goal
of being simple but not necessarily feature-rich.
The Brainstorm backend fails one more test
than the distilled background;
however, it is still an early experimental test bed,
and is not necessarily indicative of the
final Brainstorm hardware backend being
less compliant than other backends.
Finally, the SpiNNaker backend
passes the fewest tests.

It should be noted that these test suites
are being run on the distilled, Brainstorm,
and SpiNNaker backends for the first time in this study.
In other words, these results are the first objective
interrogation of each backend's feature set.
In many cases, tests failed by backends
are tests of features recently added
to Nengo, such as stochastic processes
for injecting current noise.
As such, we expect compliance on all backends
to rise quickly as backend developers
implement these new features.
The OpenCL backend has
significantly higher compliance in large part
because it is developed by the same
group that develops the reference backend.
However, there are some notable features
missing from some backends,
such as learning through plasticity rules
applied to neuron-to-neuron connections,
and implementation differences,
such as one timestep delays on connections
with no synaptic filter,
that may remain even when backends
are brought up to date with
the reference implementation.
It is likely that tests will be rewritten
in the future to allow some implementation differences
if they do not affect simulation accuracy.

@@

Figure~\ref{fig:01} gives accuracy results
for each backend on the chained communication channel model.
The boxplot shows that all five backends can
implement this model accurately,
despite five layers of processing
that each introduce noise.
The Brainstorm and distilled backends
have the least variability
and the SpiNNaker backend has the most variability,
though the median RSME is the same across all backends.
One driver of these differences
is in how each backend implement synaptic filtering.
The reference and OpenCL backends,
for example,
implement arbitrary linear time-invariant filters,
of which the lowpass filter used in this model
is a specific example.
The Brianstorm and distilled backends
only support lowpass filters,
and therefore implement it
more accurately.
Another reason for these differences
is in how each backend handles
ensembles operating near the edge
of their representational range
(i.e., the radius of the ensemble).
The outlier with highest RMSE represents
a model instance in which the static vector target
was at the extreme of the representational range;
the SpiNNaker backend's relatively high RMSE
on this example indicates that it
may not perform as well as other backends
in this situation.
However, accuracy could be improved
by increasing the representational range,
though doing so may necessitate an increase
in the number of evaluation points generated
when solving for decoding weights.

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
is less accurate than the other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.
Again, the more generic synaptic filtering model
used in the reference and OpenCL backends
is the likely cause of increased variability
on this model.
It is also likely that performance at the extremes
of the representational range
is responsible for the SpiNNaker backend's
reduced accuracy,
though we have not investigated this in detail.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
The median accuracy of the SpiNNaker is
the same as the distilled and Brainstorm backends,
but there are a high proportion of model instances
with less accuracy than the median,
as compared to other backends.
In this example, differences in synaptic filtering
are more pronounced than in previous examples,
as the recurrent connection is responsible
for the dynamics of the oscillation.
This result indicates that the synaptic filtering model
of SpiNNaker may also warrant closer examination
for accuracy.

Figure~\ref{fig:04} gives accuracy results
for the basal ganglia sequence model.
For the version of the model with passthrough nodes,
all of the backends perform similarly,
except for the SpiNNaker backend.
All other backends have a transition time
around 43 milliseconds,
but the SpiNNaker backend has a median transition time
around 51 milliseconds;
its interquartile range
is also significantly larger
than that of other backends.
However, its performance for
the version of the model
with passthrough nodes removed
is a closer match to the other backends.
Its median transition time
is around 47 milliseconds,
and its interquartile range
is indistinguishable from the other backends.
The large difference between the two versions
of this model for the SpiNNaker backend
indicates that node-to-ensemble and
ensemble-to-node connections
introduce additional delays that are
not present on other backends.
This result is to be expected,
as nodes can execute arbitrary code
and are therefore difficult
to simulate in real time
with special purpose hardware.

@@

While passthrough nodes were removed
in the basal ganglia sequence model
for improved accuracy,
it is interesting to note that speed
is also impacted in this version of the model.
While the distilled and Brainstorm backends
are unaffected,
both the SpiNNaker and OpenCL backends run faster
when passthrough nodes are removed.
This result indicates that passthrough nodes
contribute to SpiNNaker's overhead.
The reference backend, on the other hand,
runs slower,
indicating that the computational costs
of the additional connections introduced
when removing passthrough nodes
are greater than the costs of
the passthrough nodes.
